---
title: "Final Assignment"
author: "Charlotte Kuberka"
date: "2024-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(jsonlite)
library(tidyverse)
library(ggplot2)
library(sf)
library(tmap)
library(lubridate)
library(stringr)
library(dplyr)
library(ggeffects)
library(gt)
```

#### Option 2 Research question:

The R Code and additional data files can be found in the Github Repo [here](https://github.com/CharlotteKub/Final-Assignment-my472.git).

Word Count: 751 

__1. Introduction:__ 

__What, if any, characteristics and factors discriminate MPs who tend to ask questions about economic issues from MPs who tend to ask questions about health and welfare issues?__ 


To answer this research question, I used data from the UK Parliament API, and supplemented that with data on constituency characteristics. I wrote several functions to get the first 100 written and oral questions in every month for most of Theresa May’s premiership, specifically from 11/2017- 07/2019. I selected this period for its consistency under one cabinet and prime minister, avoiding skewing due to diverse governmental ideologies. I excluded the Covid period due to its health-centric focus, ensuring a balanced representation of both question topics. Thirdly, the 21-month duration allows for comparisons across different times.

__2. Managing the Data__

I merged data on written and oral questions and complemented it with more detailed MP data. To do so, I wrote another function to get data from each MP who asked an economic or health and welfare (H&W) question from the UK Parliament API and linked the datasets via the MPs name. Therefore, I had a sample with 4200 questions.


To properly identify economic and health and welfare (H&W) questions, I classified them via their answering body as in the UK parliament the questions are mostly addressed and consequently answered by the department responsible for the topic.

The classification table is shown below:


```{r echo=FALSE}

# Create the classification dataframe
classification <- data.frame(
  Economics = c('Treasury','Department for International Trade', 'HM Treasury', 'Department for Business, Energy and Industrial Strategy'), 
  Health_Welfare = c('Department of Health', 'Department of Health and Social Care', 'Women and Equalities', 'Department for Work and Pensions'))


classification <- classification %>% rename("Economic Questions" = "Economics") %>% 
  rename("Health and Welfare Questions" = "Health_Welfare")

# Create a gt table
classification_table <- classification %>%
  gt() %>%
  tab_header(
    title = "Classification Table",
    subtitle = "Which Departments were classified as Economics and Health & Welfare Departments?"
  ) 

# Print the table
classification_table

```

&ensp;

After that classification, I only included questions answered by either one of the departments described above. Thus, I had a sample with 999 question that were classified either as economic or H&W questions. I then started my analysis. To answer the RQ, I gathered MP characteristics that might influence the MPs question topic. Specifically, I decided to include gender, party, experience in parliament (seniority) and constituency characteristics in my analysis. The decision to include those factors had several reasons: firstly, party and gender have a strong effect on parliamentary activity in general (Quote). Moreover, as MPs represent a constituency it is likely that they are influenced by their characteristics in their legislative activities. Thus, MPs representing more economically deprived constituencies might ask more economic related questions than other MPs. Moreover, experience of an MP might play a role, as MPs with higher seniority tend to have more responsibility within parliament and thus ask more important questions. Economic questions could be seen as more important than H&W questions as they directly affect the budget.


```{r echo=FALSE, message=FALSE}

dates_1 <- c("2017-11-01", "2017-12-01", "2018-01-01", "2018-02-01", "2018-03-01", "2018-04-01", "2018-05-01", "2018-06-01", "2018-07-01", "2018-08-01", "2018-09-01", "2018-10-01", "2018-11-01", "2018-12-01", "2019-01-01", "2019-02-01", "2019-03-01", "2019-04-01", "2019-05-01", "2019-06-01", "2019-07-01")
dates_2 <- c("2017-12-01", "2018-01-01", "2018-02-01", "2018-03-01", "2018-04-01", "2018-05-01", "2018-06-01", "2018-07-01", "2018-08-01", "2018-09-01", "2018-10-01", "2018-11-01", "2018-12-01", "2019-01-01", "2019-02-01", "2019-03-01", "2019-04-01", "2019-05-01", "2019-06-01", "2019-06-30", "2019-08-01")


written_question_url <- "https://questions-statements-api.parliament.uk/api/writtenquestions/questions?expandMember=true&take=100"


# Function for getting written data

fetch_written_data <- function(url, dates_1, dates_2) {
  result_df <- data.frame()
  
  for (i in 1:length(dates_1)) {
    # Form the API URL
    data <- fromJSON(paste0(url, "&answeredWhenFrom=", dates_1[i], "&answeredWhenTo=", dates_2[i]))
    data_response <- data$results
    
    # Append the data to the result dataframe
    result_df <- bind_rows(result_df, data_response)
  }
  return(result_df)
}


# using function to get data for written questions

written_data <- fetch_written_data(written_question_url, dates_1, dates_2)
written_data <- written_data$value
written_data$oral_written <- ("written")


# unpacking nested variable and renaming into MP_name
written_data$MP_name <- written_data$askingMember$listAs

# subsetting to merge with oral questions

written_data <- written_data %>% select(c(id,askingMemberId, MP_name, dateTabled, dateAnswered, uin, questionText,answeringBodyName,oral_written ))

```


```{r echo=FALSE, message=FALSE}
##### getting data for oral questions 

# writing function that only needs 2 parameters for oral data 

fetch_data <- function(url, dates) {
  # Initialize an empty dataframe to store the results
  result_dataframe <- data.frame()
  
  for (i in dates) {
    # Fetch data for each date
    data_oral <- fromJSON(paste0(url, i))
    
    data_results <- data_oral$Response
    
    # Append the data to the result dataframe
    result_dataframe <- bind_rows(result_dataframe, data_results)
  }
  
  return(result_dataframe)
}


## using function and getting oral data 

oral_questions_url <- "https://oralquestionsandmotions-api.parliament.uk/oralquestions/list?parameters.take=100&parameters.answeringDateStart="

oral_questions_data <- fetch_data(oral_questions_url, dates_1)
oral_questions_data$oral_written <- ("oral")

# unpacking nested variable and renaming into MP_name
oral_questions_data$MP_name <- oral_questions_data$AskingMember$ListAs

# renaming variables and subsetting oral data to merge with written questions 

oral_questions_data <- oral_questions_data %>% rename(dateTabled = TabledWhen) %>% 
  rename(askingMemberId = AskingMemberId) %>% rename(uin = UIN) %>% rename(questionText = QuestionText) %>% 
  rename(answeringBodyName = AnsweringBody) %>%
  rename(dateAnswered = AnsweringWhen) %>% rename(id = Id)

oral_questions_data <- oral_questions_data %>% select(c(id,askingMemberId, MP_name, dateTabled, dateAnswered, uin, questionText,answeringBodyName,oral_written))

```



```{r echo=FALSE, message=FALSE}
### merge oral and written questions data 
all_questions_data <- rbind(written_data, oral_questions_data)

### classify questions based on answering department and only keep those questions that are classified within economics or health and welfare

all_questions_topical <- all_questions_data  %>% filter(answeringBodyName %in% c('Treasury','Department for International Trade', 'HM Treasury','Department for Business, Energy and Industrial Strategy', 'Department of Health', 'Department of Health and Social Care', 'Women and Equalities','Department for Work and Pensions'))

all_questions_topical <- all_questions_topical %>% 
  mutate(economic_welfare = case_when(
    answeringBodyName %in% c('Treasury', 'Department for International Trade', 'HM Treasury', 'Department for Business, Energy and Industrial Strategy') ~ "economics",
    answeringBodyName %in% c('Department of Health', 'Department of Health and Social Care', 'Women and Equalities', 'Department for Work and Pensions') ~ "welfare" ))

```


```{r echo=FALSE, message=FALSE}

###############################
#### getting additional MP data
##############################

## creating vector to collect MP names who asked questions 
Askingmembers <- all_questions_topical$MP_name

members_url <- "https://members-api.parliament.uk/api/Members/Search?Name="

# Extracting the first word from each string
data_members_first_words <- str_extract(Askingmembers, "[^,\\s]+")

data_members_first_words <- unique(data_members_first_words) # attention I am not getting all with the same name


# Function 

test_data_members <-data_members_first_words[1:10]

get_members_data <- function(url, names) {
  
  data_members <-  data.frame()
  
  for (i in names) {
    member_data <- fromJSON(paste0(url, i, "&take=1&MembershipInDateRange.WasMemberOnOrAfter=2017-11-01"))
    member_data <- member_data$items
    
    data_members <- bind_rows(data_members, member_data)
  }
  
  return(data_members)
}


# using function 

members_data_final <- get_members_data(members_url, data_members_first_words)
members_data_final <- members_data_final$value
members_data_final <-members_data_final %>% rename( "askingMemberId" = "id")
```


```{r  echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
### merge members and questions data together

questions_merged <- left_join(all_questions_topical, members_data_final, by = "askingMemberId")
questions_merged <- unique(questions_merged)

# transform economic_welfare variable into factor to use it as dependent variable in my logit regression models later 

questions_merged$'Economic or Health & Welfare Question' <- as.factor(questions_merged$economic_welfare)

levels(questions_merged$'Economic or Health & Welfare Question') <- c("0", "1")

questions_merged$economic_welfare.factor <- factor(questions_merged$'Economic or Health & Welfare Question',
                                                   levels = c(0,1),
                                                   labels = c("economics", "welfare"))
```



```{r echo=FALSE, error=FALSE, warning=FALSE}
##### calculate time until answer #######
# using: dateAnswered and dateTabled

questions_merged$dateTabled <- ymd_hms(questions_merged$dateTabled)
questions_merged$dateAnswered <- ymd_hms(questions_merged$dateAnswered)
questions_merged <- questions_merged %>% mutate(answering_time = as.numeric(difftime(dateAnswered,dateTabled, units = "days")))

```

However, another important factor differentiating economic and H&W question might be the length of the answering time. To examine if there are differences between both types of questions, I calculated the difference between the date the question was tabled and the date it was answered in days. The descriptive results are shown below:

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
questions_merged %>% group_by(dateTabled, economic_welfare, oral_written) %>%
  summarize(answering_time) %>% ggplot(aes(x = dateTabled, y = answering_time, fill= economic_welfare, color = economic_welfare)) +
  geom_col() +
  scale_color_manual(values = c("lightblue","darkgreen")) +
  scale_fill_manual(values = c("lightblue", "darkgreen")) +
  facet_wrap(~ oral_written)+ 
  theme_bw() +
  ylab("Days until Answer") +
  xlab("Question ID")
```

It can be seen in the graph that there are no significant differences in the answering time between economic and H&W questions. To examine the effect of constituency characteristics on an MPs question behavior, I used data from the [House of Commons Library](https://commonslibrary.parliament.uk/constituency-data-indices-of-deprivation/) on deprivation indices for UK constituencies. I used the health and economic deprivation index with higher values indication a higher deprivation.

```{r, echo=FALSE}
#### getting data to examine deprivation of constituencies ###
setwd("~/Desktop/Data_for_Data_Scientists/Final-Assignment-my472")

deprivation <- readxl::read_excel("deprivation-dashboard.xlsx", 
                                  sheet = "Data constituencies")


# unpacked nested variable to get Constituency Name

ConstituencyName <- questions_merged$latestHouseMembership$membershipFrom

questions_merged$ConstituencyName <- ConstituencyName


# merge both datasets 

written_questions_constituencies <- left_join(questions_merged, deprivation, by = "ConstituencyName")
```


```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#### measuring seniority of MPs #####

# using membership startdate 
## create new variable with end date of government as end date of MP term

enddate_government <- c("2019-07-25T00:00:00")
written_questions_constituencies$enddate_government <- enddate_government


# unpack nested variable for parliamentary entry of MPs
parliament_entry <- written_questions_constituencies$latestHouseMembership$membershipStartDate 
written_questions_constituencies$parliament_entry <- parliament_entry

# calculating length of MP membership in days using difftime() and creating new variable called seniority 

written_questions_constituencies$parliament_entry <- ymd_hms(written_questions_constituencies$parliament_entry)
written_questions_constituencies$enddate_government <- ymd_hms(written_questions_constituencies$enddate_government)

written_questions_constituencies <- written_questions_constituencies %>% mutate (seniority = as.numeric(difftime(enddate_government,parliament_entry, units = "days")))

# rename variables name for easier use in regression models later 
written_questions_constituencies$health_index <- written_questions_constituencies$`Health deprivation and disability`
written_questions_constituencies$party <- written_questions_constituencies$latestParty$name
```

__3. Analyzing the Data__


To properly analyze the data, I will employ a logistic regression model, calculating the predicted probabilities for the MP's characteristics on asking H&W or Economic Questions.

__3.1 Logit Regression__

As already seen in the Regression Table, some of the characteristics have no (significant) effect. 

&ensp;

```{r  echo=FALSE, warning=FALSE, error=FALSE, message=FALSE,  results='asis', comment=FALSE}

labels <- c("Green Party, Ref: Conservative", "Independent, Ref: Conservative", "Labour, Ref: Conservative", "Labour(Co-op), Ref: Conservative", "Liberal Democrats, Ref: Conservative", "Independent Group, Ref: Conservative", "Gender: Male, Ref: Female", "Seniority", "Income Deprivation of Constituency", "Health Deprivation of Constituency")

DV_label = c("Health & Welfare Question (Ref: Economic Question)")

model_all <- glm( `Economic or Health & Welfare Question`~ as.factor(party) + gender + seniority + Income + health_index, data = written_questions_constituencies, family = binomial)

stargazer::stargazer(model_all, type = "html", title = "Logit Regression Model on MP characteristics influencing the type of Question",
dep.var.caption  = "Dependent Variable",
covariate.labels = labels)

```
&ensp;

__3.2 Predicted Probabilities__


However, the MP's party, gender and the constituency's deprivation seem to have an effect. Firstly, male MPs are around 20% less likely to ask H&W questions than females. Secondly, MPs from the Green party are less likely and MPs from the Labour Co-op party are more likely to ask H&W questions compared to the Conservatives as reference category. Seniority does not seem to have an effect, and the higher an MPs constituency's Health Deprivation the more they are likely to ask a H&W question and vice versa for the constituency's economic deprivation. The predicted probabilities show that trend. 

&ensp;

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=FALSE}

# calculating predicted probabilities for logit regression coefficients 

health <- ggeffects::ggemmeans(model_all, "health_index[all]")
income <- ggeffects::ggemmeans(model_all, "Income[all]")  
gender <- ggeffects::ggemmeans(model_all, "gender") 
party <- ggeffects::ggemmeans(model_all, "party") 
seniority <- ggeffects::ggemmeans(model_all, "seniority[all]") 


# Plot the predicted probabilities

health %>% ggplot(aes(x=x, y = predicted)) + geom_line() + 
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  labs(title = "Effect of MP's Constituency's Health Deprivation on Type of Question") +
  xlab("Health Deprivation Index") +
  ylab("Predicted Probability asking health & welfare question") + 
  theme_bw()

income %>% ggplot(aes(x=x, y = predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  labs(title = "Effect of MP's Constituency's Income Deprivation on Type of Question") +
  xlab("Income Deprivation Index") +
  ylab("Predicted Probability asking health & welfare question") + 
  theme_bw() 

seniority %>% ggplot(aes(x=x, y = predicted)) + geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  labs(title = "Effect of MP's Seniority on Type of Question") +
  xlab("Seniority") +
  ylab("Predicted Probability asking health & welfare question") + 
  theme_bw() 
  

plot(gender) +
  ggtitle("Effect of MP's Gender on Type of Question") +
  labs(x = "Gender", y = "Predicted Probability asking health & welfare question")

plot(party) +
  ggtitle("Effect of MP's Party on Type of Question") +
  labs(x="Party", y="Predicted Probability asking health & welfare question")

```

In a last step, I am trying to examine if there are spatial differences within the types of question. To analyze that, I used UK constituency data from the [Office for National Statistics](https://geoportal.statistics.gov.uk/search?collection=Dataset&sort=name&tags=all(BDY_PCON)). 
I merged the data with the questions data set and calculated for each constituency the number of H&W and economic questions. As displayed in the map below, it does not seem as if there are geographical differences in the types of questions. 

&ensp;


```{r echo=FALSE}
######### working with UK Geodata for constituencies #######

# load data
uk_constituencies <- sf::read_sf("UK_constituency data 2/WPC_Dec_2018_GCB_GB.shp")

# renaming variable
uk_constituencies <- uk_constituencies %>% rename(ConstituencyName = pcon18nm)

# merging data with question data 
geo_data_written_questions_constituencies <- left_join(uk_constituencies, written_questions_constituencies, by = "ConstituencyName")

# select relevant variables 
geo_data_written_questions_constituencies <- geo_data_written_questions_constituencies %>% select(c(health_index,Income, geometry, ConstituencyName, long, lat))

geo_data_written_questions_constituencies <- unique(geo_data_written_questions_constituencies)

# calculating number of economic and welfare questions for each constituency 
constituency_questions <- written_questions_constituencies %>% group_by(ConstituencyName, economic_welfare) %>%
  count()

# joining that data with the geodata 
geo_data_written_questions_constituencies <- left_join(geo_data_written_questions_constituencies, constituency_questions, by = "ConstituencyName")

# replacing NAs with 0
geo_data_written_questions_constituencies$n <- geo_data_written_questions_constituencies$n %>% replace_na(0)

# selecting relevant variables 
geo_data_written_questions_constituencies <- geo_data_written_questions_constituencies %>% select(c(geometry, economic_welfare, n, ConstituencyName, health_index, Income, long, lat ))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
tmap_mode("view")

# Attempt to fix invalid geometries
geo_data_written_questions_constituencies <- sf::st_make_valid(geo_data_written_questions_constituencies)

filtered_data <- geo_data_written_questions_constituencies %>%
  filter(!is.na(economic_welfare))

## removing id 661 as it doesnt fit the data structure for plotting (but its NA anyways)

geo_data_written_questions_constituencies <- geo_data_written_questions_constituencies %>%
  filter(ConstituencyName != 'Ross, Skye and Lochaber')


# plotting Map with UK constituencies 

tm_shape(geo_data_written_questions_constituencies) + 
  tm_polygons() +
  tm_shape(filtered_data) +
  tm_fill("n", breaks = c(0, 1, 2, 3, 4, 5, 10, 20), palette = "Blues", title = "Questions No.") +
  tm_facets(by = "economic_welfare", free.coords = TRUE, drop.empty.facets = TRUE) +
   tm_layout(
    title = c("Economic Questions", "Health & Welfare Questions"))

```


&ensp;

__4. Conclusion__


In general, my analysis showed that factors such as gender, party and constituency characteristics of MPs discriminate wether or not MPs ask economic or H&W questions. However, future research could try and analyze if there are characteristics in the answers given to those questions that discriminate by topic.

#### Appendix: All code in this assignment

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 
# this chunk generates the complete code appendix. 
# eval=FALSE tells R not to run (``evaluate'') the code here (it was already run before).
```
